---
output: html_document
---
PML Project.
---

It was requested to analize data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and create the model whch would be able predict the manner in which they did the exercise *classe* variable:

Data was available on internet. I get it and save on my project folder in file with the same name *pml-training.csv*. First my step was to read the data:
```{r message=FALSE} 
dataset <- read.csv("pml-training.csv") 
```


Preprocessing steps.
--

Data file contains 19622 observations.
On the next preprocessing step 1 several columns will be removed: 
- which contain 19216 *NA* values;
- which most freequent value is empty space.
```{r message=FALSE} 
summ <- summary(dataset);
get_freequent_value <- function(x)
{
 return (names(table(x))[1])
}
empty <- which(apply(dataset, 2, get_freequent_value) == "")
NA_s1 <- which(summ[7,] == "NA's   :19216  ")
remove <- c(empty, NA_s1)
dataset <- dataset[, -remove]
```
After preprocessing 60 variables left.


Now we should review what is left. Get index of outcome variable, order variable and character variables which should be removed from prediction dataset in future.  
```{r message=FALSE}
# summary(dataset);
# names(dataset); head(dataset)
outcome <- which(names(dataset) == c("X", "classe"))
charact <- c(1, 2, 5, 6)
```


It's time for preprocessing step 2 on which we would be able find related columns. We will use correlation koefficient and search for columns which are linearly related.  
```{r message=FALSE}
relation_check <- dataset[,-c(charact, outcome)]
M <- abs(cor(relation_check))
diag(M) <- 0
related_ <- which(M > 0.8, arr.ind=T)
related_ <- related_[which(related_[,1] < related_[,2]),]
related_ <- related_[order(related_[,1]),]
related <- unique(related_[,2])
related <- related[order(related)]

lower_numbers <- function(x, charact)
{
  num = 0
  charact[order(charact)]
  for(y in charact){
    if (x >= y)
    {
      x = x + 1
      num = num + 1
    }
  }
  return (num)
}
related <- related + sapply(related, lower_numbers, charact)
```


Data analysis
--

Here ispart to analize relations beetween variables.
# To DO: more information to investigate the difference and wrong / right behaviour.
```{r message=FALSE}
dataset <- dataset[,-c(charact, related)]
# pairs(training[,c(1:9,47)])
```

Here is time to split data to Train and Test sets.
```{r message=FALSE}
suppressWarnings(library(caret));
inTrain <- createDataPartition(y=dataset$classe, p=0.7, list=FALSE)
training <- dataset[inTrain,]; testing <- dataset[-inTrain,]
```


Model training
--

Here we will train *Random Forest* model, the best from the tested. 
```{r message=FALSE}
suppressWarnings(library(randomForest));
suppressWarnings(library(foreach));
rf <- foreach(n=rep(200, 5), .combine=randomForest::combine, .packages='randomForest') %do% {
  randomForest(training[, -length(training)], training$classe, ntree=n)
}
modelFit <- predict(rf, newdata=training)
pred <- predict(rf, newdata=testing) 
table(pred, testing$classe)  #confusionMatrix(pred, testing$classe)
```
Model accuracy 0.999.

 


Use model in practice.
--
As I have best result for model *Random Forest*. I'll use it in test case prediction. 
```{r message=FALSE}
dataset_test <- read.csv("pml-testing.csv") 
dataset_test <- dataset_test[, -remove]
dataset_test <- dataset_test[, -c(charact, related)]  # related

pred <- predict(rf, newdata=dataset_test[,-length(dataset_test)]);
pred
```


Results.
--

*Random Forest* model was used to evaluate relation.